A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-09-21 docs/icccn2009/paper.tex
Replaced former DSP EPS images with simpler ones.
git-svn-id: svn://localhost/netbeams/trunk@421 559edde2-a4d4-44e0-96fc-4035e40d7396
Resubmitting paper with DSP diagrams using Open Office. Both the new EPS and original ODG files have been uploaded.
git-svn-id: svn://localhost/netbeams/trunk@422 559edde2-a4d4-44e0-96fc-4035e40d7396
A very tiny edit.
git-svn-id: svn://localhost/netbeams/trunk@423 559edde2-a4d4-44e0-96fc-4035e40d7396
REVIEWS DONE BY KLEBER... I'M JUST COMMITTING HIS CHANGES DESCRIBED AT REVISION 77 OF THE PAPER REVISION GOOGLE DOCUMENT.
http://docs.google.com/Revs?id=0Af5ISdu-g4

A wireless sensor network consists of spatially distributed autonomous
devices using sensors to cooperatively monitor physical or
environmental conditions at different locations \cite{roemer:2004}.
Much research effort is focused on building wireless ad-hoc networks, 
where each sensor node participates in a multi-hop routing algorithm.
One of the earliest examples of a wireless sensor network is the
SmartDust project \cite{smartdust:2001}, where one objective was to
create autonomous sensing and communication in a cubic
millimeter-sized device.  While most wireless sensor networks focus on
small sensing devices such as the SmartDust mote, there exist a wide
variety of off-the-shelf commercially available sensors that are used
by environmental researchers. In contrast to the small sensing
devices, sensors for environmental monitoring are typically on the
order of one to two feet in length, weighing several pounds. Our
focus is on building an end-to-end infrastructure for environmental
monitoring using specialized sensing devices.

@@

NetBEAMS's architecture is based on a micro-kernel approach. The core
functionality is limited to a minimum and all extra services are built
as plug-and-play modules on top of the micro-kernel. With the help of
this approach, we can keep the basic infrastructure compact while
allowing customization via special-purpose modules. In our terminology,
we call the basic foundation the \emph{Data Sensor Platform}
(DSP). The scope of the DSP is one address space of an execution
platform and a sensor network is built from several DSPs that are linked
with each other.

@@

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP. Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 6) and stopped (line
7) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
8). A DP can send messages via a \texttt{DSPContext} object that was
passed to the DP during initialization.

@@

\begin{figure}[!htb]
 \centering
 \epsfig{file=dsp-message, width=9cm, height=4cm}
 \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP message.}
\end{figure}

The basic message structure is defined by an \texttt{AbstractMessage}
type, whose structure serves as a base schema for all different types
of messages. Any message will contain two basic attributes:
\texttt{messageID} and \texttt{ContentType}.  The former uniquely
identifies a message and the latter the data in the body of the
message. Complementary identification information, such as the message
producer and consumer, is located in the header of the message and 
is defined by the schema type \texttt{ComponentIdentifier}. A
\texttt{ComponentIdentifier} carries out information about the
component name and its physical host address.  Finally, the body of
the message represents the payload and may contain any data type
defined by the component developer.

@@

\begin{figure}[!htb]
 \epsfig{file=dsp-match-rule, width=9cm, height=4cm}
 \caption{\label{FIG_MATCHER_RULE} Definition of Matcher Rule.}
\end{figure}

@@

The DSP is implemented in Java and serves as a container environment where 
DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics, 
the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
infrastucture. One important DSP requirement is to impose on DSP Components minimal 
dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
accessible to DSP Components are specified by \texttt{DSPContext} interface.
The DSP also uses messages to communicate with DSP Components. It has a special adapter
that acts like a \texttt{DSPComponent}. Therefore, configuring the DSP is no different from configuring
a DSP Component.

@@

\begin{figure}[!htb]
\centering
\epsfig{file=dsp-message-body-ysi, width=9cm}
\caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
\end{figure}

>>>>

A wireless sensor network consists of spatially distributed autonomous
devices using sensors to cooperatively monitor physical or
environmental conditions at different locations \cite{roemer:2004}.
Much research effort is focused on building wireless ad-hoc networks, 
where each sensor node participates in a multi-hop routing algorithm.
One of the earliest examples of a wireless sensor network is the
SmartDust project \cite{smartdust:2001}, where one objective was to
create autonomous sensing and communication in a cubic
millimeter-sized device.  While most wireless sensor networks focus on
small sensing devices such as the SmartDust mote, there exists a wide
variety of off-the-shelf commercially available sensors that are used
by environmental researchers. In contrast to the small sensing
devices, sensors for environmental monitoring are typically on the
order of one to two feet in length, weighing several pounds. Our
focus is on building an end-to-end infrastructure for environmental
monitoring using specialized sensing devices.

@@

NetBEAMS's architecture is based on a micro-kernel approach. The core
functionality is limited to a minimum and all extra services are built
as plug-and-play modules on top of the micro-kernel. With the help of
this approach, we can keep the basic infrastructure compact while
allowing customization via special-purpose modules. In our terminology,
we call the basic foundation the \emph{Data Sensor Platform}
(DSP). DSP is collection of local built-in and plug-in components; 
a sensor network is built from several DSPs that are linked
with each other.

@@

By implementing the above Java interface, the resulting code
effectively becomes a DSP Component that can be deployed on a DSP at
runtime. A component is identified by its type (line 3). Only one
component of a certain type per DSP is permissible. A type identifier
must be unique across all DSP Components in a DSP. Upon deploying a
component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 6) and stopped (line
7) multiple times. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (line
8). A DP can send messages via a \texttt{DSP Broker} (a DSP Component itself) 
obtained from \texttt{DSPContext} object that was
passed to the DP during initialization.

@@

\begin{figure}[!htb]
 \centering
 \epsfig{file=AbstractMessage2, width=9cm, height=4cm}
 \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP message.}
\end{figure}

The basic message structure is defined by an \texttt{AbstractMessage}
type, whose structure serves as a base schema for all different types
of messages. Any message will contain two basic attributes:
\texttt{messageID} and \texttt{contentType}.  The former uniquely
identifies a message and the latter the data in the body of the
message. Complementary identification information, such as the message
producer and consumer, is located in the header of the message and 
is defined by the schema type \texttt{ComponentIdentifier}. A
\texttt{ComponentIdentifier} carries out information about the
component name and its physical host address.  Finally, the body of
the message represents the payload and may contain any data type
defined by the component developer.

@@

\begin{figure}[!htb]
 \centering
 \epsfig{file=MatcherRule2, width=8cm, height=5cm}
 \caption{\label{FIG_MATCHER_RULE} Definition of Matcher Rule.}
\end{figure}

@@

The DSP is implemented in Java and serves as a container environment where 
DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics, 
the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
infrastucture. One important DSP requirement is to impose on DSP Components minimal 
dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
accessible to DSP Components are specified by \texttt{DSPContext} interface.
The DSP communicates with DSP Components through the same message mechanism DSP components
used among themselves. Therefore, configuring the DSP is no different from configuring
a DSP Component, which simplifies administration and monitoring implementations.

@@

\begin{figure}[!htb]
\centering
\epsfig{file=YSI_Message_Format2, width=9cm}
\caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
\end{figure}
